---
title: "Computational Musicology Portfolio 2025"
author: "Xuelong Fu"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    theme:
      version: 4
      # #ffa2ff
      bg: "#c0fcf8"
      fg: "#005b5c" 
      primary: "#47817d"
      navbar-bg: "#3ADAC6"
      base_font: 
        google: Shippori Mincho
      heading_font:
        google: Shippori Mincho
---

```{r setup, include=FALSE}
library(flexdashboard)
library(readr)
library(ggplot2)
library(dplyr)
library(lubridate)
library(ggrepel)
library(sysfonts)
library(showtext)
library(tidyverse)
source("compmus.R")

font_add_google("Shippori Mincho", "Shippori Mincho")
showtext_auto()
```

```{r load_data, include = FALSE}
df <- read_csv('../data/compmus2025.csv')
```

TEMP Homework Week 11
=====================================

Histogram
-----------------------------------------------------------------------

### Class Corpus Histogram

```{r}
ggplot(df, aes(x = tempo)) +
  geom_histogram(binwidth = 5, fill = "#47817d", color = "white", alpha = 0.8) +
  labs(title = "Tempo Distribution", x = "Tempo", y = "Frequency") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    panel.background = element_rect(fill = "#c0fcf8", color = NA),
    plot.background = element_rect(fill = "#c0fcf8", color = NA)
  )
```

Tempograms 
-----------------------------------------------------------------------

We notice that the tempo curves are quite constant, where little surprising is 
happening. In the first tempo curve, we do see some sort of interference at 80-90,
just like in the self-similarity matrix. This shows that a section of different detected 
tempo arrangement and timbre is happening there in my first track. This is probably due
as indeed my track changes to something very different there, where dubstep sounds take
over what is generally Drum n Bass in that section. 

### Track 1

```{r}
"../data/features/xuelong-f-1.json" |>
  compmus_tempogram(window_size = 8, hop_size = 1, cyclic = FALSE) |>
  ggplot(aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()
```

### Track 2

```{r}
"../data/features/xuelong-f-2.json" |>
  compmus_tempogram(window_size = 8, hop_size = 1, cyclic = FALSE) |>
  ggplot(aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()
```


TEMP Homework Week 9
=====================================

Chromagram
-----------------------------------------------------------------------

### Track 1 - Chromagram

```{r}
"../data/features/xuelong-f-1.json" |>               
  compmus_chroma(norm = "identity") |>                
  ggplot(aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c(
                "C", "C#|Db", "D", "D#|Eb",
                "E", "F", "F#|Gb", "G",
                "G#|Ab", "A", "A#|Bb", "B"
              )
  ) +
  scale_fill_viridis_c(guide = "none") +            
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic() 
```

### Track 2 - Chromagram

```{r}
"../data/features/xuelong-f-2.json" |>                        
  compmus_chroma(norm = "identity") |>              
  ggplot(aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c(
                "C", "C#|Db", "D", "D#|Eb",
                "E", "F", "F#|Gb", "G",
                "G#|Ab", "A", "A#|Bb", "B"
              )
  ) +
  scale_fill_viridis_c(guide = "none") +              
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic() 
```

Ceptrogram
-----------------------------------------------------------------------

### Track 1 - Ceptrogram

```{r}
"../data/features/xuelong-f-1.json" |>                      
  compmus_mfccs(norm = "identity") |>          
  ggplot(aes(x = time, y = mfcc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:12,
    minor_breaks = NULL,
  ) +
  scale_fill_viridis_c(guide = "none") +            
  labs(x = "Time (s)", y = "Coefficient Number", fill = NULL) +
  theme_classic()   
```

### Track 2 - Ceptrogram

```{r}
"../data/features/xuelong-f-2.json" |>                        
  compmus_mfccs(norm = "identity") |>               
  ggplot(aes(x = time, y = mfcc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:12,
    minor_breaks = NULL,
  ) +
  scale_fill_viridis_c(guide = "none") +              
  labs(x = "Time (s)", y = "Coefficient Number", fill = NULL) +
  theme_classic() 
```

Timbre-based self-similarity
-----------------------------------------------------------------------

(Aesthetics and design elements for all the graphs will be done later). We notice that our tracks are reasonably clearly structured, with limited variation in certain sections. There are limited changes in timbral palettes, textures, and instrumentation. And as you listen to them, it indeed kind of makes sense. When I made these tracks, I indeed attempted to make them more structured (e.g. with clearly delineated song sections) and the shifts in instrumentation and timbre was indeed quite limited. As feedback for myself, this may make my tracks a bit stale over time. So it is great to gain insight in these kind of aspects, as they may help me in my future music production endeavors. 

### Track 1 - Timbre-based self-similarity

```{r}
"../data/features/xuelong-f-1.json" |>                           # Change the track
  compmus_mfccs(norm = "identity") |>                  # Change the norm
  compmus_self_similarity(
    feature = mfcc,
    distance = "euclidean"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic() 
```


### Track 2 - Timbre-based self-similarity

```{r}
"../data/features/xuelong-f-2.json" |>                           # Change the track
  compmus_mfccs(norm = "identity") |>                  # Change the norm
  compmus_self_similarity(
    feature = mfcc,
    distance = "euclidean"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic() 
```

Welcome
=====================================

Description
-----------------------------------------------------------------------

For this homework, I used my two tracks I submitted previously. Here are the descriptions:

#### Track xuelong-f-1

This track is a remix/arrangement by me of a videogame OST song. The original is the "Stage 5" theme called "Faraway Voyage of 380,000 Kilometers" (https://youtu.be/a8Wdta9YFYk?si=mZuIxx1y-nJusEIQ) in the game "Touhou 15, Legacy of Lunatic Kingdom". It is composed by ZUN and has lenient copyright requirements (especially if it is not commercialized). For my remix, I used elements of the original composition and made a Drum-n-Bass track of it. Listen it here:

<audio controls preload="none">
  <source src="../data/xuelong-f-1.mp3" type="audio/mpeg">
  Your browser does not support the audio element.
</audio>

#### Track xuelong-f-2

This track is a remix/arrangement by me of an OST song of the same game franchise. The original is the "Stage 1 Boss" theme called "Kitten of Great Fortune" (https://youtu.be/yeleqYd0t5o?si=KWyjivretXP-fEP3) in the game "Touhou 18, Unconnected Marketeers". It is, again, composed by ZUN and has lenient copyright requirements. For my remix, I used elements of the original composition and made a Electro Swing track of it. Listen it here:

<audio controls preload="none">
  <source src="../data/xuelong-f-2.mp3" type="audio/mpeg">
  Your browser does not support the audio element.
</audio>

One of my goals (perhaps) for the portfolio is to compare musical elements, like timbre and richness of sound, between my remixes and the original tracks.

TEMP Homework Week 8
=====================================

K-Means Clustering
-----------------------------------------------------------------------

### K-Means Clustering

```{r}
features <- df %>% 
  select(-filename)

features_scaled <- scale(features)

set.seed(42)
kmeans_result <- kmeans(features_scaled, centers = 3)

df <- df %>%
  mutate(cluster = as.factor(kmeans_result$cluster))

pca <- prcomp(features_scaled, center = TRUE, scale. = FALSE)
pca_data <- data.frame(PC1 = pca$x[, 1],
                       PC2 = pca$x[, 2],
                       filename = df$filename,
                       cluster = df$cluster)

custom_theme <- theme_minimal(base_family = "Shippori Mincho") +
  theme(
    plot.background = element_rect(fill = "#c0fcf8", color = NA),
    panel.background = element_rect(fill = "#c0fcf8", color = NA),
    text = element_text(color = "#005b5c"),
    plot.title = element_text(family = "Shippori Mincho", size = 16, face = "bold", color = "#005b5c"),
    axis.title = element_text(family = "Shippori Mincho", size = 12, color = "#005b5c"),
    legend.background = element_rect(fill = "#3ADAC6", color = NA),
    legend.key = element_rect(fill = "#101010", color = NA),
    panel.grid.minor = element_blank()
  )

cluster_colors <- c("#005b5c", "#6a417a", "#6b4b11")

cluster_plot_1 <- ggplot(pca_data, aes(x = PC1, y = PC2, color = cluster, label = filename)) +
  geom_point(size = 1.5) +
  geom_point(data = subset(pca_data, filename %in% c("xuelong-f-1", "xuelong-f-2")),
             aes(x = PC1, y = PC2),
             color = "#ff8748", size = 2.5) +
  geom_text_repel(size = 3.3, family = "Shippori Mincho", max.overlaps = Inf) +
  labs(title = "Clustering Analysis based on All Features",
       x = "Principal Component 1",
       y = "Principal Component 2",
       color = "Cluster") +
  scale_color_manual(values = cluster_colors) +
  custom_theme

print(cluster_plot_1)
```

Description
-----------------------------------------------------------------------

Generally, my tracks are classified as more "energetic" compared to the class corpus. For instance, they score high on danceability, engagingness, and valence. Based on the clustering analysis, we observe that the Essentia features describe my own and other students' tracks quite reasonably. We note that my tracks are generally in the third cluster, where more of the energetic tracks reside, such as aleksandra-b-1 or mette-l-2. My tracks are positioned further away from tracks like jasper-v-1, which is a more relaxing jazz-like track, or lo-l-1, which is a more calm orchestral song. In this way, we notice that the features generally make sense, although some mistakes remain such as that Essentia provided wrong BPM tempos for my track. It detected the energetic nature of my tracks and how it relates and fits besides the other energetic tracks in the corpus. 


Exploratory Analysis
=====================================

Here I aim to perform a general overview of the data

Feature Analysis
=====================================

Here I aim to analyze features as generated by Essentia

Deep Dive
=====================================

Here I aim to dive deeper into specific songs (probably the boss themes of Touhou Project)

Comparisons
=====================================

Here I aim to perform comparisons, such as originals vs. my remixes of the originals

Final Remarks
=====================================
